---
title: "final paper"
output: html_document
date: "2025-01-08" 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


question: whether the democracy have effect on economic growth 

data explanation

GDP per capita (GROWTH), per capita: per person

Inflation (INF), 

Trade (TRADE), 

final Government consumption (GOVTSIZE), 

Population growth (POP), 

Political stability (PST), Values range between -2.5 and +2.5, with -2.5 being the most unstable situation and +2.5 indicating the most stable: measure the quality of governance


Deliberative democracy index (DELIBDEM), The democratic deliberative principle
concerns how decisions are made in society.



```{r}
# import dataset
df <- read.csv('/Users/akane/Documents/applied econometric/final assignment/Panel_democ_stability_growth_MENA_Over_1983_2022.csv')
head(df)
library(AER)
library(plm)
library(stargazer)
```



```{r}
# the set of observations on all variables for 2012 and 2022
df_1997To2022 <- subset(df, df$year %in% c(1997:2022))
describe <- df_1997To2022[,c(5,6,7,8,9,10,12)]
stargazer(describe, type = "latex")


```


## The Fixed Effects Regression Mode

Regression using panel data may mitigate omitted variable bias, which could be a feasible choice as the complex impact that have no information on variables may exist when it comes to policy issues. 


First, to The simple fixed effects regression model for estimation of the relation between economic growth and the democracy is

$$
GROWTH_{it} = \beta_1 DEM_{it} + CountryFixedEffects + u_{it}
$$

a regression of the democracy on economic growth and 15 binary regressors-- one for each MENA country.


```{r}

# estimate the fixed effects regression with plm()
# the ID variable for index is "country" and "year"
#fixed effects estimator is achieved by setting model = "within"
gdp_dem_mod2 <- plm(GROWTH ~ delibdem,
data = df_1997To2022,
index = c("country", "year"),
model = "within")
# partipdem delibdem egaldem
# print summary usingg robust standard errors
coeftest(gdp_dem_mod2, vcov. = vcovHC, type = "HC1")

```

The estimated coefficient is -11.4192, which is negative and sufficient.

The interpretation is that the estimated reduction in economic growth due to an increase in democracy by 1% is 11 per capita, which actually shows a conflicting views on the effect of democracy on growth.

For the definition of the model, obtain inference based on robust standard errors is crucial, which means that clustered standard errors is applied for fixed effects models.


we plot the data and add the corresponding estimated regression functions. 




## The Fixed Effects Regression Assumptions and Standard Errors for Fixed Effects Regression

There are several major source of omitted variable bias that are not accounted for by all of the models of the relationship between economic growth and democracy measurement.These covariates are defined as follows:


Inflation (INF), 
Trade (TRADE), 
final Government consumption (GOVTSIZE),
Population growth (POP) 
Political stability (PST). 


Thus the model with all of the considering covariate is following: 

$$
\text { GROWTH }_{i t}=\alpha_i + \beta_1 \mathrm{INF}_{\mathrm{it}} +\beta_2 \text { TRADE }_{\mathrm{it}}+\beta_3 \mathrm{GOVTSIZE}_{\mathrm{it}}+\beta_4 \mathrm{POP}_{\mathrm{it}} +\beta_5 \mathrm{PST}_{\mathrm{it}}+ \beta_6 \mathrm{DEM}_{\mathrm{it}} + \varepsilon_{i t}
$$ 
where $\alpha_i$ is a separate constant term for each of the i MENA country. 




## Estimate several models and compare them. 



```{r}
# estimate all seven models
gdp_dem_mod1 <- lm(GROWTH ~ delibdem, data = df_1997To2022)
# summary(gdp_dem_mod1) 


# gdp_dem_mod2 <- plm(GROWTH ~ delibdem,
# data = df_1997To2022,
# index = c("country", "year"),
# model = "within")


gdp_dem_mod3 <- plm(GROWTH ~ INF + TRADE  + delibdem ,
data = result,
index = c("country", "year_group"),
model = "within"
)
coeftest(gdp_dem_mod3, vcov = vcovHC, type = "HC1")

# add variables which is related to government and society 
gdp_dem_mod4 <- plm(GROWTH ~ GROWTH_lag + INF + POP + delibdem + TRADE + GOVTSIZE  + PST, 
                   data = result,
                   index = c("country", "year"),
                   model = "within"
)
summary(gdp_dem_mod4)

# print summary using robust standard errors

# coeftest(gdp_dem_mod4, vcov = vcovHC, type = "HC1") 
```


```{r}

# gather clustered standard errors in a list
rob_se <- list(sqrt(diag(vcovHC(gdp_dem_mod1, type = "HC1"))),
               sqrt(diag(vcovHC(gdp_dem_mod2, type = "HC1"))),
               sqrt(diag(vcovHC(gdp_dem_mod3, type = "HC1"))),
               sqrt(diag(vcovHC(gdp_dem_mod4, type = "HC1"))))
# generate the table
stargazer(gdp_dem_mod1, gdp_dem_mod2, gdp_dem_mod3,
gdp_dem_mod4, 
digits = 3,
header = FALSE,
type = "latex",
se = rob_se,
title = "Linear Panel Regression Models of Democracy Effect on Economic Growth",
model.numbers = FALSE,
column.labels = c("(1)", "(2)", "(3)", "(4)"))



```











# Spatial dynamic panel data modeling

```{r}
library(SDPDmod)
library(rnaturalearth)
library(sf)
```


# calculating the mean value for every 5 years.


```{r}
df$year_group <- cut(df$year, 
                     breaks = seq(floor(min(df$year) / 5) * 5, ceiling(max(df$year) / 5) * 5, by = 5),
                     include.lowest = TRUE, 
                     right = FALSE)

library(dplyr)

result <- df %>%
  group_by(country, year_group) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE), .groups = "drop")

```


```{r}
result <- data.frame(result)

library(Hmisc)

result$GROWTH_lag <- c(NA, result$GROWTH[-nrow(result)])
result$GROWTH_lag[which(!duplicated(result$country))] <- NA
result$GROWTH_lag = impute(result$GROWTH_lag,0)
head(result)

d2 <- plm::pdata.frame(result, index=c('country', 'year_group'))
d2$llogc<-plm::lag(d2$logc) 
head(d2)
```

```{r}
d2 <- plm::pdata.frame(result, index=c('country', 'year'))
```


```{r}
mena <- unique(df$country)

# Read in the world countries shapefile
world <- ne_countries(scale = "medium", returnclass = "sf")
mena_world <- world[world$name %in% mena, ]


mena_world <- mena_world[match(mena, mena_world$name),] # match countries by the order of original data. 

ggplot(mena_world) +
  geom_sf() +
  geom_sf_text(data = mena_world, aes(label = name))
plot(mena_world["name"], main = "MENA Countries")
```


```{r}
# Spatial Contiguity Weights Matrix
W_2n_norm <- mOrdNbr(sf_pol = mena_world, m = 2, rn = T) ## second order neighbors
W_len_sh <- SharedBMat(mena_world,rn = T)
W_len_sh_norm <- rownor(W_len_sh)


```




```{r}

library(splm)

usalw <- spdep::mat2listw(W_len_sh_norm,zero.policy = T)



sar_model <- spml(GROWTH ~ INF, 
                  data = result,
                  index = c("country", "year_group"),
                  model = "within",
                  listw = usalw, 
                  zero.policy = T,
                  lag=TRUE)


a <- summary(sar_model)

library(texreg)
texreg(a, type = "latex")

stargazer(sar_model, 
digits = 3,
header = FALSE,
type = "text",
title = " Democracy Effect on Economic Growth",
model.numbers = FALSE)
```










```{r}
# Static spatial lag model with spatial fixed effects (ignore)
mod1<-SDPDm(formula = GROWTH ~ INF + POP + delibdem + TRADE + GOVTSIZE + PST, data =result, W = W_len_sh_norm,
            index = c("country","year"), 
            model = "sar", 
            effect = "individual")


summary(mod1)
```

```{r}
# Dynamic spatial model with spatial fixed effects
mod2<-SDPDm(formula = GROWTH ~ INF + POP + delibdem + TRADE + GOVTSIZE  + PST, data = data.frame(result), W = W_len_sh_norm,
            index = c("country","year"),
            model = "sar", 
            effect = "individual",
            dynamic = T, 
            tlaginfo = list(ind = NULL, tl = T, stl = T))

summary(mod2)


```


```{r}
# Dynamic spatial Durbin model with time fixed effects
mod3<-SDPDm(formula = GROWTH ~ INF + POP + delibdem + TRADE + GOVTSIZE  + PST, data = data.frame(result), W = W_len_sh_norm,
            index = c("country","year"),
            model = "sdm", 
            effect = "individual",
            dynamic = T,
            tlaginfo = list(ind = NULL, tl = T, stl = T))


```


```{r}
# Extract coefficients, standard errors, and p-values
model_summary <- data.frame(summary(gdp_dem_mod4)$coefficients)

coefficients <- model_summary$Estimate
std_errors <- model_summary$Std..Error
p_values <- model_summary$Pr...t..
#coefficients <- c(model_summary$coefficients,model_summary$rho )  # Coefficients
#std_errors <- c(model_summary$std, model_summary$rho.se)  # Standard Errors
#p_values <- c(model_summary$pval,model_summary$rho.pval)  # P-values

# Create a custom function to format p-values with asterisks
format_p_values <- function(p) {
  if (p < 0.01) {
    return("***")
  } else if (p < 0.05) {
    return("**")
  } else if (p < 0.1) {
    return("*")
  } else {
    return("")
  }
}

# Format the coefficients, standard errors, and p-values for LaTeX
formatted_results <- paste0(round(coefficients, 3), sapply(p_values, format_p_values))


# Create a data frame for the results
results_table <- data.frame(
  Variable = rownames(model_summary),
  Coefficient = formatted_results
)

# Render the table to LaTeX 
latex_table <- xtable(results_table, caption = "Regression Results with Blank Column Name")
print(latex_table, type = "latex", include.rownames = FALSE)

```





```{r}
# calculating impacts
imp  <- impactsSDPDm(mod5)
summary(imp)
```

```{r}

# Extract direct, indirect, and total impacts
direct <- imp
indirect <- imp$INDIRECTst.tab
total <- imp$TOTALst.tab

# Combine into a data frame
impact_results <- data.frame(
  Variable = rownames(direct),
  Direct = direct[, 1],  # Extract point estimates for direct impacts
  Indirect = indirect[, 1],  # Extract point estimates for indirect impacts
  Total = total[, 1]  # Extract point estimates for total impacts
)




```



```{r}
#Dynamic panel model with both spatial and time fixed effects with uniform prior
res4<-blmpSDPD(formula = GROWTH ~ INF + POP + delibdem + TRADE + GOVTSIZE  + PST, data = data.frame(result), W = W_len_sh_norm, 
            index = c("country","year"),
               model = list("ols","sar","sdm"), 
               effect = "individual",
               ldet = "mc", ## log-determinant calculated with mcmc procedure
               dynamic = TRUE,
               prior = "uniform")

# Static panel model with  log-marginal 
res2<-blmpSDPD(formula = GROWTH ~ INF + POP + delibdem + TRADE + GOVTSIZE  + PST + GROWTH_lag, data = data.frame(result), W = W_len_sh_norm,
               index = c("country","year"),
               model = list("ols","sar","sdm"), 
               effect = "individual")

print(res2)


```


